---
title: "COMPSCIX 415.2 Homework 7"
author: "Michelle Gomez"
date: "7/18/2018"
output:
  pdf_document: default
  html_document: default
self_contained: yes
toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mdsr)
library(broom)
```
#Exercise 1
```{r fig.height=10, message=FALSE, warning=FALSE}
train <- file.path("/Users/michellegomez/Downloads/train.csv")
train_data <- read_csv(train)
glimpse(train_data)
```

#Exercise 2  
```{r}
train_data %>%
  group_by(SalePrice) %>%
  ggplot() + geom_histogram(aes(x = SalePrice))

train_data %>%
  select(SalePrice, Neighborhood) %>% arrange_all() %>% ggplot() + geom_boxplot(aes(x = Neighborhood, y = SalePrice)) + coord_flip()

train_data %>%
  select(SalePrice, OverallQual) %>% arrange_all() %>% ggplot() + geom_smooth(aes(x = OverallQual, y = SalePrice)) + coord_flip()

```

#Exercise 3  
```{r}
(sale_lm <- lm(formula = SalePrice ~ OverallQual, data = train_data))
tidy(sale_lm)
glance(sale_lm)
mean(train_data$SalePrice)
45435.80/180921.2
```
- take a look at the coefficient  
The coefficient is 45435.80 increase in SalePrice by unit increase in OverallQual.  
- compare the coefficient to the average value of SalePrice  
The coefficient seems to be 1/4 of the average value of Sale Price.
- take a look at the R-squared  
R-squaredis .62, which can be considered a good fit.  

#Exercise 4  
```{r}
sale_mult_lm <- lm(formula = SalePrice ~ GrLivArea + OverallQual + Neighborhood, data = train_data)
sale_mult_lm
tidy(sale_mult_lm)
glance(sale_mult_lm)
```

- How would you interpret the coefficients on GrLivArea and OverallQual?  
For every one unit increase in GrLivArea, the SalePrice increases, on average, by $55.57.  
For every one unit increase in OverallQual, the SalePrice increases, on average, by $20951.43.  
- How would you interpret the coefficient on NeighborhoodBrkSide?  
The mean SalePrice difference between NeighborhoodBrkSide and Blmngtn is -$13025.45.  
- Are the features significant?  
The p-values are all less than alpha=.05 therefore they are all significant prefictors of SalePrice.  

- Are the features practically significant?  
I don't think that the features are practically siginificant because you can't increase them by a unit, rather they describe a relationship of something measurable like price to some more subjective features.  
- Is the model a good fit?  
The model's adjusted R-squared is 0.78, which is closer to 1 so it's more a less a "good" fit but could definitely be better.  

#Exercise 6
One downside of the linear model is that it is sensitive to unusual values because the distance incorporates a squared term. Fit a linear model to the simulated data below (use y as the target and x as the feature), and look at the resulting coefficients and R-squared. Rerun it about 5-6 times to generate different simulated datasets. What do you notice about the model’s coefficient on x and the R-squared values?  
```{r}
sim1a <- tibble(
x = rep(1:10, each = 3),
y = x * 1.5 + 6 + rt(length(x), df = 2)
)
ggplot(sim1a) + geom_point(aes(x = x, y = y))
sim1a_lm <- lm(formula = y ~ x , data = sim1a)
tidy(sim1a_lm)
glance(sim1a_lm)
```
I notice that the R-squared value keeps jumping up and down between .06-.99 and the x coefficient is also variable from 1.0-1.5. I think that the linear model can be improved by weighing outliers less than values that fall within range.


